Tree establishment in models of forest dynamics: a key priority for further research

### 2022-12-15 manuscript/manuscriptRW.Rmd
references updates

---
title: "Tree establishment in models of forest dynamics: a key priority for further research"
author:
- Olalla Díaz-Yáñez,
- Yannek Käber,
- Tim Anders,
- Kristin H. Braziunas,
- Josef	Bruna,
- Samuel Fischer,
- Jessica Hetzer,
- Thomas Hickler,
- Christian	Hochauer,
- Manfred Lexer,
- Heike	Lischke,
- Mats Mahnken,
- Paola	Mairota,
- Katarina Merganicova,
- Tobias Mette,
- Xavier Morin,
- Werner Rammer,
- Simon	Scheiter,
- Daniel Scherrer,
- Harald Bugmann
date: "`r Sys.Date()`"
output:
  bookdown::pdf_document2:
    keep_tex: yes
    number_sections: yes
    toc: no
  word_document:
    reference_docx: template_paper.docx
mainfont: Arial
fontsize: 12pt
linestretch: 1.5
bibliography: Regeneration.bib
link-citations: yes
nocite: |
  @reyer, @bugmann1996a, @bohn2014, @fabrika2005, @mette2014, @smith2001a, @seidl2012a, @lischke2006a, @schumacher2006, @scheiter2009, @scheller2004, @morin2021
linkcolor: blue
csl: global-change-biology.csl
editor_options:
  markdown:
    wrap: 72
header-includes:
- \usepackage[left]{lineno}
- \linenumbers
---

@@

Thus, a focus on the modeling of tree establishment processes is sorely
needed and overdue [@price2001] if we are to make reliable projections
of future forest dynamics, i.e., when the models need to be operated in
extrapolation mode, but also from a fundamental ecological point of view
(increasing systems understanding). Also in the field of general
ecology, tree establishment has not received a lot of attention
[@hanburybrown_forest_2022], although many sub-processes have been
studied in detail. Yet, little data is available that covers all the
processes within one species along environmental gradients, let alone
for a vast suite of species. Thus, the available data are fragmented,
which constitutes a major problem for model building. On the one hand,
monitoring on permanent sample plots (such as National Forest
Inventories) often measure tree establishment [@zell2019], however with
a design that captures the rate after the stand initiation phase
[@hallsby2015]. Data from permanent plots, although available across
regions, are highly heterogeneous, rendering their utility difficult in
a modeling context[@käber2021]. On the other hand, targeted studies to
measure tree establishment on experimental sites are often limited in
spatial extent, species studied, or the sub-set of processes that are
targeted [@berdanier2016; @collet2006].

@@

The design of our sampling protocol did not include spatial aspects. We
considered a wide range of models of forest dynamics, from stand to
global scale. However, the simulation setup was limited to 1 ha ingrowth
values. The lack of spatial considerations potentially puts landscape
models at a disadvantage, as they have been built to be accurate at the
landscape level. Without spatial context, we are limiting ingrowth to
the seed influx from the stand itself, if the model does not have a
background seed input. However, Global Dynamic Vegetation Models should
not be in disadvantage due to the limited spatial considerations [but
cf. Snell et al. -@snell2014] as they are usually lacking dispersal
between cells as they are based on an abstraction of horizontal space,
or even the trees themselves [@hanburybrown_forest_2022].

>>>>

---
title: "Tree establishment in models of forest dynamics: a key priority for further research"
author:
- Olalla Díaz-Yáñez,
- Yannek Käber,
- Tim Anders,
- Kristin H. Braziunas,
- Josef	Bruna,
- Samuel Fischer,
- Jessica Hetzer,
- Thomas Hickler,
- Christian	Hochauer,
- Manfred Lexer,
- Heike	Lischke,
- Mats Mahnken,
- Paola	Mairota,
- Katarina Merganicova,
- Tobias Mette,
- Xavier Morin,
- Werner Rammer,
- Simon	Scheiter,
- Daniel Scherrer,
- Harald Bugmann
date: "`r Sys.Date()`"
output:
  bookdown::pdf_document2:
    keep_tex: yes
    number_sections: yes
    toc: no
  word_document:
    reference_docx: template_paper.docx
mainfont: Arial
fontsize: 12pt
linestretch: 1.5
bibliography: Regeneration.bib
link-citations: yes
nocite: |
  @reyer, @bugmann1996a, @bohn2014, @fabrika2005, @mette2014, @smith2001a, @seidl2012a, @lischke2006, @schumacher2006, @scheiter2009, @scheller2004, @morin2021
linkcolor: blue
csl: global-change-biology.csl
editor_options:
  markdown:
    wrap: 72
header-includes:
- \usepackage[left]{lineno}
- \linenumbers
---

@@

Thus, a focus on the modeling of tree establishment processes is sorely
needed and overdue [@price2001] if we are to make reliable projections
of future forest dynamics, i.e., when the models need to be operated in
extrapolation mode, but also from a fundamental ecological point of view
(increasing systems understanding). Also in the field of general
ecology, tree establishment has not received a lot of attention
[@hanbury-brown2022], although many sub-processes have been studied in
detail. Yet, little data is available that covers all the processes
within one species along environmental gradients, let alone for a vast
suite of species. Thus, the available data are fragmented, which
constitutes a major problem for model building. On the one hand,
monitoring on permanent sample plots (such as National Forest
Inventories) often measure tree establishment [@zell2019], however with
a design that captures the rate after the stand initiation phase
[@hallsby2015]. Data from permanent plots, although available across
regions, are highly heterogeneous, rendering their utility difficult in
a modeling context [@käber2021]. On the other hand, targeted studies to
measure tree establishment on experimental sites are often limited in
spatial extent, species studied, or the sub-set of processes that are
targeted [@berdanier2016; @collet2006].

@@

The design of our sampling protocol did not include spatial aspects. We
considered a wide range of models of forest dynamics, from stand to
global scale. However, the simulation setup was limited to 1 ha ingrowth
values. The lack of spatial considerations potentially puts landscape
models at a disadvantage, as they have been built to be accurate at the
landscape level. Without spatial context, we are limiting ingrowth to
the seed influx from the stand itself, if the model does not have a
background seed input. However, Global Dynamic Vegetation Models should
not be in disadvantage due to the limited spatial considerations [but
cf. Snell et al. -@snell2014] as they are usually lacking dispersal
between cells as they are based on an abstraction of horizontal space,
or even the trees themselves [@hanbury-brown2022].
